---
title: "RT-qPCR 2026-01-18 on SUN4-mNG 5' UTR mutant strains"
author: "Benny Kleinerman & Edward Wallace"
date: "18/01/2026"
output:
  html_document:
    toc: yes
    toc_depth: 2
---

# Experimental Design

Replicates:

- 3 biological replicates are collected for each strain
- 3 technical replicates +RT and 1 -RT

Test Primers:

- mNG_1
- SUN4_1

Reference gene primers: 
- ACT1_1
- PGK1_2
- SRO9_1
- SSD1_1

```{r load_tidyqpcr}
library(tidyqpcr)
library(forcats)
library(ggplot2)
library(dplyr)
library(tidyr)
library(readr)
library(ggbeeswarm)
```

# Describe the experimental design in a plate plan

## Columns are target_ids of primer sets

```{r target_ids}
colkey <- create_colkey_6_in_24(
  target_id = c("mNG_1", "SUN4_1", "ACT1_1", "PGK1_2", "SRO9_1", "SSD1_1") %>% 
    as_factor()
)

print(colkey)
```


## Rows are sample name, Strain, BioRep (3) and TechRep (3)

Also choose colours to display strains.

```{r samples_1-8}
strain_df <- tibble(
  Strain = c(
    "yMJ003",
    "yBK023",
    "yBK012",
    "yMJ011",
    "yBK032",
    "yBK013",
    "yMJ039",
    "yBK010",
    "yMJ041",
    "yBK011"),
  fiveprimeUTR = c(
    "WT",
    "intronless",
    "PGK1",
    "WT",
    "intronless",
    "PGK1",
    "WT",
    "PGK1",
    "WT",
    "PGK1"),
  threeprimeUTR = c(rep("ADH1", 6), rep("SUN4", 4)),
  background = c(
    rep("Wild Type", 3),
    rep("ssd1delta", 3),
    rep("Wild Type", 2),
    rep("ssd1delta", 2)),
  colour = rgb(runif(10), runif(10), runif(10)))

scale_color_strain <- 
  scale_color_manual("Strain",
                     breaks = strain_df$Strain,
                     values = strain_df$colour)

rowkey <- tibble(well_row = c((LETTERS[1:16]), LETTERS[1:14]), 
                 Strain = strain_df$Strain %>% 
                   rep(each = 3) %>% 
                   as_factor(),
                 fiveprimeUTR = strain_df$fiveprimeUTR %>% 
                   rep(each = 3) %>% 
                   as_factor(),
                 threeprimeUTR = strain_df$threeprimeUTR %>% 
                   rep(each = 3) %>% 
                   as_factor(),
                 background = strain_df$background %>% 
                   rep(each = 3) %>% 
                   as_factor(),
                 BioRep = 1:3 %>% 
                   as_factor() %>%
                   rep(times = 10)) %>%
  unite(sample_id, Strain, BioRep, remove = FALSE) %>%
  mutate(sample_id = as_factor(sample_id))

print(rowkey)
```

## Combine into plate plan

```{r plate_plan_1, fig.width = 12, fig.height = 8}

plate_plan_1 <- label_plate_rowcol(
  plate = create_blank_plate(well_row = LETTERS[1:16], well_col = 1:24),
  rowkey = rowkey %>%
  dplyr::slice(1:16),
  colkey = colkey
)

print(plate_plan_1)
display_plate_qpcr(plate_plan_1)

```

```{r plate_plan_2, fig.width = 12, fig.height = 8}

plate_plan_2 <- label_plate_rowcol(
  plate = create_blank_plate(well_row = LETTERS[1:14], well_col = 1:24),
  rowkey = rowkey %>%
  dplyr::slice(17:30),
  colkey = colkey
)

print(plate_plan_2)
display_plate_qpcr(plate_plan_2)

```

# Load data of RT-qPCR results in Cq format

```{r load_data}

plate_1 <- here::here("2026-01-28_SUN4-mNG_RTqPCR_replacedUTRs/2026-01-28_SUN4-mNG_RTqPCR_replacedUTRs_p1_BK_Ct.txt") %>%
  read_tsv(skip=1) %>%
    mutate(well=Pos,cq=Cp, plate_id = "1") %>%
    right_join(plate_plan_1)

plate_2 <- here::here("2026-01-28_SUN4-mNG_RTqPCR_replacedUTRs/2026-01-28_SUN4-mNG_RTqPCR_replacedUTRs_p2_BK_Ct.txt") %>%
  read_tsv(skip=1) %>%
    mutate(well=Pos,cq=Cp, plate_id = "2") %>%
    right_join(plate_plan_2)

plates <- bind_rows(plate_1, plate_2)
plates

```

# Plot all replicates, unnormalized

```{r plot_plate, fig.width = 7, fig.height = 12}

ggplot(data = plates) +
  geom_point(aes(x = target_id, y = cq, shape = prep_type, colour = target_id),
    position = position_jitter(width = 0.2, height = 0)
  ) +
  labs(
    y = "Quantification cycle (Cq)",
    title = "All reps, unnormalized, plate"
  ) +
  facet_wrap(~sample_id, ncol = 5) +
  theme_classic() + 
  theme(panel.grid.major.y = element_line(colour= "grey80"),
        panel.grid.minor = element_blank(),
        axis.line = element_blank(),
        axis.ticks = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5))

```

The high or absent Cq values in -RT reactions indicate absence of DNA contamination.


# Double-check Cq values of SUN4-mNeonGreen scarless in WT and ssd1delta backgrounds

```{r plot_plate_SUN4-mNG-scarless, fig.width = 6, fig.height = 7}

ggplot(data = plates %>%
         dplyr::filter(Strain %in% c("yMJ039", "yMJ041"))) +
  geom_point(aes(x = target_id, y = cq, shape = prep_type, colour = target_id),
    position = position_jitter(width = 0.2, height = 0)
  ) +
  labs(
    y = "Quantification cycle (Cq)",
    title = "All reps, unnormalized, plate"
  ) +
  facet_wrap(~sample_id, ncol = 3) +
  scale_y_continuous(limits = c(12,20), breaks = seq(10,25)) +
  theme_classic() + 
  theme(panel.grid.major.y = element_line(colour= "grey80"),
        panel.grid.minor = element_blank(),
        axis.line = element_blank(),
        axis.ticks = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5))

```

ssd1delta (yMJ041) mNG Cq values are higher than WT. This is opposite to the effect of ssd1 deletion on Sun4-mNeonGreen protein expression

# Check reference genes have consistent Cq across samples

```{r plate_sep_refgenes, fig.width = 12, fig.height = 4}
ggplot(data = dplyr::filter(plates, target_id %in% c("ACT1_1", "PGK1_2", "SRO9_1") )) +
  geom_point(aes(x = target_id, y = cq, shape = prep_type, colour = target_id),
    position = position_jitter(width = 0.2, height = 0)) +
  facet_grid(~sample_id) +
  theme_classic() + 
  theme(panel.grid.major.y = element_line(colour= "grey80"),
        panel.grid.minor = element_blank(),
        axis.line = element_blank(),
        axis.ticks = element_blank(),
        axis.text.x = element_blank(),
        strip.text.x = element_text(angle = 90, vjust = 0.5))
```

Reference genes have fairly consistent Cq across samples

# Normalise data by calculating ∆Cq against median of 3 reference genes

```{r plate_norm}

plate_norm <- plates %>%
  dplyr::filter(prep_type == "+RT") %>%
  droplevels() %>%
  calculate_deltacq_bysampleid(ref_target_ids = c("ACT1_1", "PGK1_2", "SRO9_1"))

plate_norm

plate_norm_med <- plate_norm %>%
  group_by(sample_id, Strain, fiveprimeUTR, threeprimeUTR, background, BioRep, target_id) %>%
  summarize(
    delta_cq = median(delta_cq, na.rm = TRUE),
    rel_abund = median(rel_abund, na.rm = TRUE)
  )

plate_norm_med

```

# Plot ∆Cq for normalized data

For all replicates of all genes, all samples.

```{r plot_deltacq_all, fig.width = 8, fig.height = 4}

ggplot(data = plate_norm) +
  geom_point(aes(x = target_id, y = delta_cq, shape = BioRep),
    position = position_jitter(width = 0.2, height = 0)
  ) +
  labs(x = "Target Gene", 
         y = "Relative mRNA detection (∆Cq)") +
  theme_classic() + 
  theme(panel.grid.major.y = element_line(colour= "grey80"),
        panel.grid.minor = element_blank(),
        axis.line = element_blank(),
        axis.ticks = element_blank(),
        axis.text.x = element_text(size = 10, angle = 90))

```

DeltaCq values are similar across strains and replicates

# Normalise data by calculating ∆∆Cq against median of 1 reference sample (yMJ003)

```{r plate_deltanorm}
plate_deltanorm <- plate_norm %>%
  calculate_deltadeltacq_bytargetid(ref_sample_ids = c("yMJ003_1", "yMJ003_2", "yMJ003_3"))

plate_deltanorm_med <- plate_deltanorm %>%
    mutate(UTRs = interaction(fiveprimeUTR, threeprimeUTR) %>%
                  forcats::fct_recode(
                    "PGK1 ADH1" = "PGK1.ADH1",
                    "SUN4 (no intron) ADH1" = "intronless.ADH1",
                    "SUN4 ADH1" = "WT.ADH1",
                    "PGK1 SUN4" = "PGK1.SUN4",
                    "SUN4 SUN4" = "WT.SUN4") %>%
                  forcats::fct_relevel(
                    "PGK1 ADH1",
                    "SUN4 (no intron) ADH1",
                    "SUN4 ADH1",
                    "PGK1 SUN4",
                    "SUN4 SUN4")) %>%
    group_by(sample_id, target_id, Strain, fiveprimeUTR, threeprimeUTR, UTRs, background, BioRep) %>%
  summarize(
    deltadelta_cq  = median(deltadelta_cq, na.rm = TRUE),
    fold_change    = median(fold_change,   na.rm = TRUE)
  )

plate_deltanorm_med
```

# Plot ∆∆Cq for normalized data

For all genes, all samples, summarized by biological replicate.

```{r plot_deltadelta_cq, fig.width = 8, fig.height = 4}
set.seed(4)
ggplot(data = plate_deltanorm_med,
       aes(x = target_id, y = deltadelta_cq)
       ) +
  geom_point(
    position = position_jitter(width = 0.4, height = 0),
    size = 2
  ) +
  stat_summary(geom = "crossbar", fun = mean, fun.min = NULL, fun.max = NULL, aes(group=1)) +
  labs(x = "Target Gene", 
       y = "∆∆Cq (log2 fold-change)") +
  theme_classic() + 
  theme(panel.grid.major.y = element_line(colour= "grey80"),
        panel.grid.minor = element_blank(),
        axis.line = element_blank(),
        axis.ticks = element_blank(),
        axis.text.x = element_text(size = 10, angle = 90))

```

deltadeltaCq values are close to 0, for target and reference genes.

# Plot deltadeltaCq values for mNeonGreen across all samples

```{r deltadeltacq_mNG_utrs, fig.width = 6, fig.height = 7}
mNG <- ggplot(data = dplyr::filter(plate_deltanorm_med, target_id %in% "mNG_1")) +
  geom_point(aes(x = deltadelta_cq, y = UTRs,
                 colour = background,
                 group = interaction(Strain,BioRep),
                 shape = BioRep,
                 size = 2,
                 stroke = 2),
    position = position_quasirandom(groupOnX = FALSE, width = 0.15)) +
  theme_classic() + 
  labs(x = "ΔΔCq", y = "Strain") +
  theme(panel.grid.major.x = element_line(colour= "grey80"),
        panel.grid.minor = element_blank(),
        axis.line = element_blank(),
        axis.text.y = element_text(size = 14),
        axis.text.x = element_text(size = 10),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 14),
        axis.ticks = element_blank()) +
        scale_x_continuous(limits = c(-1.5, 1.5),
                     breaks = c(-1, -.5, 0, .5, 1),
                     labels = c("-1", "-0.5", "0", "0.5", "1")) +
  guides(colour = guide_legend(override.aes = list(linetype = c(0, 0), fill = NA, alpha = 1)), size = "none") + 
  scale_colour_manual(values = c("#005a00ff", "#7edc7eff")) +
  scale_shape_manual(values = c(0,1,2))

mNG
```

# Calculate linear model coefficients, to compare against flow cytometry signal intensities

```{r lm_deltadeltacq_mNG_utrs, fig.width = 6, fig.height = 7} 
plate_deltanorm_med_coeff <- plate_deltanorm_med %>% 
          mutate(Strain = interaction(fiveprimeUTR, threeprimeUTR, background)) %>%
          mutate(Strain = relevel(Strain, ref = "WT.SUN4.Wild Type"))

fit_RTqPCR_utrs <- lm(deltadelta_cq ~ 0 + Strain, # removes intercept
   data = dplyr::filter(plate_deltanorm_med_coeff, target_id %in% "mNG_1"))

# make a table from the linear model coefficients
fit_RTqPCR_utrs_tsv <- as.data.frame(summary(fit_RTqPCR_utrs)$coefficients)

# write the table out to a .tsv file
write.table(
  fit_RTqPCR_utrs_tsv,
  file = "fit_RTqPCR_utrs.tsv",
  sep = "\t",
  quote = FALSE,
  col.names = NA)
```

# Fit a linear model that uses the no intron reference strain as intercept, and calculate fold changes

```{r lm_deltadeltacq_mNG_utrs_2, fig.width = 6, fig.height = 7} 
fit_RTqPCR_utrs_diffs <- lm(deltadelta_cq ~ Strain, # includes intercept
   data = dplyr::filter(plate_deltanorm_med_coeff, target_id %in% "mNG_1"))

foldchange_RTqPCR_utrs_diffs <- 2^{ coef(fit_RTqPCR_utrs_diffs)[-1] }
foldchange_RTqPCR_utrs_diffs
```

