---
title: "2026-05-11_Ssd1_pointmutant_flow"
author: "Benny Kleinerman"
output:
  html_document:
    toc: yes
    toc_depth: 2
date: "2026-05-11"
---

# Summary

Summary: Analysis of the flow cytometry data gathered 2026-02-17, on episomal Ssd1 RNA-binding point mutants, in a Sun4-mNeonGreen reporter background.

This script only includes code to generate figures for the paper. Additional code is in 2026-02-17_Ssd1_pm_redo or similar scripts.

# Load packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, messages = FALSE, warning = FALSE)
library(tidyverse)
library(cowplot)
theme_set(theme_cowplot(font_size = 12) +
            theme(strip.background = element_blank()))
library(ggridges)
library(flowCore)
library(ggbeeswarm)
library(ggpubr)
```

# Define functions to load flow cytometry data, using FlowCore package

```{r functions_load}
repair_flow_colnames <- function(string) {
  string %>%
    stringr::str_replace_all(" ", "") %>%
    stringr::str_replace_all("PE-", "PE") %>%
    stringr::str_replace_all("-", "_")
}

read_fcs_tibble <- function(file,...) {
  read.FCS(file,...) %>%
    .@exprs %>%
    as_tibble %>%
    set_names(repair_flow_colnames(names(.)))
}
```

# Describe the data

First we load the sample sheet, that describes the sample contents and filenames.

```{r load_samplesheet}
datadir <- here::here("2026-02-17_Ssd1_pm_redo")

samplesheet <- paste(datadir, "/2026-02-17_Ssd1_pm_redo_plan.xlsx", sep = "") %>%
  readxl::read_excel(skip = 4) %>%
  mutate(Biorep = factor(Biorep))

print(samplesheet, n = 19)
```

# Load all the data

Next we load the data from all those files.

```{r load_all_data}
flow_data_all <- samplesheet  %>%
  group_by(Sample, ID, Biorep, SSD1_homolog_unstructured, SSD1_homolog_structured, Expression, mRuby2, Culture_media, Strain, Experiment, Filename) %>%
  do( read_fcs_tibble(paste(datadir, .$Filename[1], sep = "/")) )
```

# Gate data

Here we gate the data for budded singlets. We have identified singlets in spectral flow cytometry analysis from February 2026

Gates for other populations are in the older scripts, described in the Summary

```{r gate_data}
# Gate containing budded singlet cells, as identified by imaging flow cytometry
flow_data_gated_buddedsinglets <- flow_data_all %>%
  dplyr::filter(FSC_H > 30000, FSC_H < 60000,
                FSC_W > 66000, FSC_W < 72000)
```

# How many events remain after gating?

```{r inspect_flow_data}
flow_data_all

flow_count_gates <- 
  left_join(flow_data_all %>%
              group_by(Filename) %>%
              tally(),
            flow_data_gated_buddedsinglets %>%
              group_by(Filename) %>%
              tally(),
            by = c("Filename"),
            suffix = c("_all", "_gated") ) %>%
  mutate(prop_gated = n_gated / n_all)

print(flow_count_gates, n = 19)
```

Samples still have thousands of events after gating

# Compare mNeonGreen (FITC) signal with FSC_A (forward scatter area)

## mRuby2 control (ssd1Δ)

```{r FITC_vs_FSC_A_ssd1Δ, fig.height = 3, fig.width = 4}
ggplot(
  data = dplyr::filter(flow_data_gated_buddedsinglets, Strain == "dual stain"),
  aes(x = FSC_A, y = `530_30(488)_A`)) +
  geom_hex(bins = 64) +
  scale_fill_gradientn(colours = rainbow(6)[1:5]) +
  stat_cor(
    method = "pearson",
    aes(label = ..r.label..),
    label.sep = "",      
    r.accuracy = 0.01,    
    label.x = 30000,
    label.y = 12000) +
  ylim(c(0, 12000))
```

## WT

```{r FITC_vs_FSC_A_WT, fig.height = 3, fig.width = 4}
ggplot(
  data = dplyr::filter(flow_data_gated_buddedsinglets, Strain == "pRNR2-ScSSD1-mRuby2"),
  aes(x = FSC_A, y = `530_30(488)_A`)) +
  geom_hex(bins = 64) +
  scale_fill_gradientn(colours = rainbow(6)[1:5]) +
  stat_cor(
    method = "pearson",
    aes(label = ..r.label..),
    label.sep = "",      
    r.accuracy = 0.01,    
    label.x = 30000,
    label.y = 6000) +
  ylim(c(0, 7500))
```

mNeonGreen scales with FSC_A, at least for these 2 strains

# Ridgeplots of mNeonGreen for ssd1 point mutant strains

## First create a new column in the budded singlets dataset, with readable Strain names

```{r ssd1pm_ssd1locus}
flow_data_gated_buddedsinglets <- flow_data_gated_buddedsinglets %>%
  mutate(ssd1locus = forcats::fct_recode(Strain,
      "WT" = "pRNR2-ScSSD1-mRuby2",
      "Empty" = "ssd1Δ",
      "Empty mRuby2" = "dual stain",
      "R394A" = "pRNR2-Scssd1R394A-mRuby2",
      "W583A" = "pRNR2-Scssd1W583A-mRuby2",
      "K589A" = "pRNR2-Scssd1K589A-mRuby2",
      "KR622-623AA" = "pRNR2-Scssd1KR622-623AA-mRuby2",
      "H630A" = "pRNR2-Scssd1H630A-mRuby2"),
    ssd1locus = fct_rev(ssd1locus))
```

## SUN4-mNeonGreen, with FSC normalization

```{r ssd1pm_FITC/FSC_ridgeplot, fig.height =5, fig.width = 8}

mNG <- ggplot(
  data = flow_data_gated_buddedsinglets,
  aes(x = `530_30(488)_A` / FSC_A, y = ssd1locus)) +
  stat_density_ridges(
    scale = 1,
    linewidth = 1.5,
    alpha = 0.2,
    colour = "#008000ff",
    fill = "#008000ff",
    bandwidth = 0.01) +
  stat_summary(
    aes(group = interaction(Strain, Biorep), shape = Biorep),
    colour = "#008000ff",
    fun = mean,
    geom = "point",
    size = 4,
    stroke = 2,
    position = position_quasirandom(groupOnX = FALSE, width = 0.1)) +
  labs(x = "mNeonGreen/FSC-A", y = "Strain") +
  theme(
    panel.grid.major.x = element_line(colour = "grey80"),
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    axis.text.y = element_text(vjust = -0.1)) +
  coord_cartesian(expand = TRUE, clip = 'off') +
  scale_x_continuous(limits = c(0, 0.2)) +
  scale_shape_manual(values = c(0, 1, 2))

mNG

# I manually annotate axis labels and legends in inkscape

ggsave(
  filename = "U:/data/2026/05-May/Benny/2026-05-11_Ssd1_pointmutants_mNG_flow.svg",
  plot = mNG,
  device = "svg",
  width = 10,
  height = 10,
  limitsize = FALSE)
```

Generally, RNA-binding Ssd1 point mutants fail to regulate Sun4-mNeonGreen expression. The exception to this is the K589A mutant

## Ssd1-mRuby2

```{r ssd1pm_mRuby2_ridgeplot_small, fig.height = 5, fig.width = 8}

mRuby2 <- ggplot(dplyr::filter(
  flow_data_gated_buddedsinglets, `610_20(561)_A` > 0),
  aes(x = `610_20(561)_A`, y = ssd1locus)) +
  stat_density_ridges(
    scale = 1,
    linewidth = 1.5,
    alpha = 0.2,
    bandwidth = 0.15,
    colour = "#C48A95",
    fill = "#F6C1D1") +
  stat_summary(
    aes(group = interaction(Strain, Biorep), shape = Biorep),
    fun = mean,
    geom = "point",
    size = 4,
    stroke = 2,
    colour = "#C48A95",
    position = position_quasirandom(groupOnX = FALSE, width = 0.1)) +
  labs(x = "mRuby2", y = "Strain") +
  theme(
    panel.grid.major.x = element_line(colour = "grey80"),
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    axis.text.y = element_text(vjust = -0.1)) +
  coord_cartesian(expand = TRUE, clip = 'off') +
  scale_x_log10(breaks = c(10, 1000, 100000),
    labels = c("10", "1000", "100000")) +
  scale_colour_manual(values = c("#C48A95")) +
  scale_fill_manual(values = c("#F6C1D1")) +
  scale_shape_manual(values = c(0, 1, 2))

mRuby2

# I manually annotate axis labels and legends in inkscape

ggsave(
  filename = "U:/data/2026/05-May/Benny/2026-03-14_Ssd1_pointmutant_mRuby2_flow.svg",
  plot = mRuby2,
  device = "svg",
  width = 10,
  height = 10,
  limitsize = FALSE)
```

All point mutants are expressing from plasmids under pRNR2. Translational repression aligns well with growth on CFW

# Calculate fold changes between the strain means

## mRuby2, comparing all strains to the empty plasmid strain

```{r ssd1pm_mRuby2_fc}
# Set Empty plasmid strain as the reference

flow_data_gated_buddedsinglets_emptyref <- flow_data_gated_buddedsinglets %>%
  dplyr::filter(`610_20(561)_A` > 0) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(ssd1locus = relevel(as.factor(ssd1locus), ref = "Empty"))

# log10 linear model

fit_mRuby2 <- lm(
  log10(`610_20(561)_A`) ~ ssd1locus,
  data = flow_data_gated_buddedsinglets_emptyref)

foldchange_mRuby2 <- 10^{ coef(fit_mRuby2)[-1] }
foldchange_mRuby2
```

The empty mRuby2 construct (no Ssd1) produces 35 times as much red signal as background in an mRuby2-less strain, and the constructs with Ssd1 produce ~10 times as much red signal

## mNeonGreen, comparing all strains to wild-type Ssd1

```{r ssd1pm_mNG/FSC_fc}
# Set WT Ssd1 strain as the reference

flow_data_gated_buddedsinglets_WTref <- flow_data_gated_buddedsinglets %>%
  dplyr::filter(`530_30(488)_A` > 0) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(ssd1locus = relevel(as.factor(ssd1locus), ref = "WT"))

# linear model

fit_mNG <- lm(
  log10(`530_30(488)_A`/FSC_A) ~ ssd1locus,
  data = flow_data_gated_buddedsinglets_WTref)

foldchange_mNG <- 10^{ coef(fit_mNG)[-1] }
foldchange_mNG
```
There is 2.72 times as much mNeonGreen in ssd1 deletion cells as Wild Type Ssd1